% <- percent signs are used to comment
\documentclass[12pt]{article}

%%%%%% PACKAGES - this part loads additional material for LaTeX %%%%%%%%%
% Nearly anything you want can be done in LaTeX if you load the right package 
% (search ctan.org or google it if you are looking for something).  We will load
% here a few that we need for this document or that we expect you to need later.

% The next 3 lines are needed to fix shortcomings of TeX that only make sense given its 40-year history ...
% Simple keep and ignore.
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath}
\usepackage{changepage}
\usepackage{lipsum}

% Custom margins (and paper sizes etc.) because LaTeX else wastes much space
\usepackage[margin=1in]{geometry}

% The following packages are created by the American Mathematical Society (AMS)
% and provide lots of tools for special fonts, symbols, theorems, and proof
\usepackage{amsmath,amsfonts,amssymb,amsthm}
% mathtools contains many detail improvements over ams and core tex
\usepackage{mathtools}

% graphicx is required for images
\usepackage{graphicx}

% enumitem used for customizing enumerations
\usepackage[shortlabels]{enumitem}

% tikz is the package used for drawing, in particular for drawing trees. You may also find simplified packages like tikz-qtree and forest useful
\usepackage{tikz}

% hyperref allows links, urls, and many other PDF tricks.  We load it here
%          in such a way that the PDF file has info about it
\usepackage[%
	pdftitle={CS240 Assignment 0},%
	hidelinks,%
]{hyperref}


%%%%%% COMMANDS - here you can define your own LaTeX-commands %%%%%%%%%

%%%%%% End of Preamble %%%%%%%%%%%%%

\begin{document}

\begin{center}
{\Large\textbf{CS240, Spring 2022}}\\
\vspace{2mm}
{\Large\textbf{Assignment 2: Question 2}}\\
\vspace{3mm}
\end{center}
\[ \]
\textbf{Q2)} We want to prove the following: there is no comparison-based algorithm that can merge $m$ sorted arrays of length $m$ into a unique sorted
array of length $m^2$ doing $O(m^2)$ comparisons. We argue by
contradiction, and we assume that it is possible, so that we have such
an algorithm (which we call FastMerge).

Modify MergeSort in order to use FastMerge, and derive a
contradiction. You may use the following property: if a function
$T(n)$ satisfies $T(n) = \sqrt{n} T(\sqrt{n}) + O(n)$, then $T(n) =
O(n \log(\log(n)))$. Do not worry about $n$ being a perfect square or
not.\\
\begin{adjustwidth}{0em}{0pt}
We will prove using contradiction, that is we will assume there is a comparison-based algorithm (called FastMerge) that can merge $m$ sorted arrays of length $m$ into a unique sorted array of length $m^2$ doing $O(m^2)$ comparisons. \\ \\ Moving on we will modify merge sort so that it will do the following steps:
\begin{enumerate}
  \item Split the starting array into $\sqrt{n}$ new arrays of size $\sqrt{n}$ and sort theses new arrays.
  \item Merged them back into a bigger arrays using FastMerge.
\end{enumerate} Step (1) will involve $\sqrt{n}$ recursive calls all of which have a time complexity of:
\[ T(n) = \sqrt{n}T(\sqrt{n}) \]
And since splitting happens in constant time, we wont need to factor it in.\\
Step (2) will take $O(n)$ time as we assumed when starting our proof. Thus we will have a total time complexity of:
\[ T(n) = \sqrt{n}T(\sqrt{n}) + O(n) \]
From what we are given in the question we get the new equation:
\[ T(n) = O(n \log(\log(n))) \]
However we know from the slides that merge sort has a time complexity of:
\[ T(n) = O(n \log(n)) \]
In order to compare these time complexities we will do a limit comparison test to get:
\begin{align*}
    \begin{aligned}
       L &= \lim_{n->\infty}\frac{n \log(\log(n))}{n \log(n)}  \\
       &= \lim_{n->\infty}\frac{\log(\log(n))}{\log(n)} \\
    \end{aligned}
\end{align*}
Before moving forward we will make a variable $x=\log(n)$ and then take the derivative in respect to $x$, note that as $n$ approaches $\infty$ so will $x$.
\begin{align*}
    \begin{aligned}
       L &= \lim_{x->\infty}\frac{\log(x)}{x}  \\
       &= \lim_{x->\infty}\frac{\frac{1}{x}x - ln(x)}{x^2} \\
       &= \lim_{x->\infty}\frac{1 - ln(x)}{x^2} \\
       &= \lim_{x->\infty}\frac{1}{x^2} - \frac{ln(x)}{x^2}  \\
       &= \lim_{x->\infty}\frac{1}{x^2} - \frac{\frac{ln(x)}{x}}{x}  \\
       &= \frac{1}{\infty} - \frac{0}{\infty}  \\
       &= 0 \\
    \end{aligned}
\end{align*}
Therefore by the limit test we get that: 
\[ n \log(\log(n)) \in o(n \log(n)) \]
This means that $n \log(\log(n))$ has a growth rate less then $n \log(n)$, which is a contradiction because we know that merge sort has a time complexity of $O(n \log(n))$.
Thus we have disproved by contradiction that there is no comparison-based algorithm that can merge $m$ sorted arrays of length $m$ into a unique sorted
array of length $m^2$ doing $O(m^2)$ comparisons. 
\end{adjustwidth} 




\end{document}